---
title: "African Elephant Populations"
author: "Camryn Allen"
date: March 27, 2024
output: 
  beamer_presentation:
    theme: "Malmoe" 
    colortheme: "beaver"
fontsize: 11pt
---

```{r, eval=FALSE, echo=FALSE} 
knitr::opts_chunk$set( #what does eval & echo = FALSE mean? 
  echo = FALSE,
  message = FALSE, #meaning?
  warning = FALSE,#meaning?
  prompt = FALSE, #meaning? 
  tidy = FALSE, #meaning?
  comment = NA, #meaning? 
  fig.align = 'center'
  )

library(tidyverse)
library(knitr)
library(rmdformats)
library(readr)
library(ggplot2)
library(patchwork)
```

# The Data Details

Collected by the African Elephant Specialist Group (AfESG) of the Food and Agriculture Organization of the United Nations, the data utilized for this presentation visualizes and compares the elephant populations of different African countries at intervals between 1979 and 2015.


Overview

- Africa's Elephant Population (1979 - 2015)
- Elephant Populations Across 28 Countries (2015)


# A Focus on Africa's Elephant Population Overtime
```{r, echo=FALSE, eval=TRUE, warning=FALSE,message=FALSE, prompt=FALSE, comment=NA} 
library(dplyr)
library(readr)
african_elephants <- read_csv('african-elephants.csv')

df2 = african_elephants

AfricaElephants = 
  african_elephants %>%
  filter(Entity %in% c('Africa'))

AfricaElephants <- subset(AfricaElephants, Year > 1913)

AfrMostElephants = 
  african_elephants %>%
  filter(Entity %in% c('Botswana', 'Tanzania', 'Zimbabwe'))

AfrMostElephants <- AfrMostElephants %>%
  filter(Year != '2007')

#Plot of African Elephants
library(ggplot2)
AfricaPlot = 
  AfricaElephants |>
  ggplot(aes(x = Year, y = `African Elephant population (AfESG, 2019)`)) + 
  geom_line(
    color = '#74A9CE'
  ) + 
  ggtitle('African Elephant Population (1979 - 2015)') + 
  labs(
    x = 'Year',
    y = "Elephant Population"
  ) +
  scale_x_continuous(
    breaks = c(1970,1980,1990,2000,2010)
  ) + 
  scale_y_continuous(
    labels = c('250,000', '500,000', '750,000', '1.0M', '12.5M')
  ) +
  theme_bw() + 
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = 'bold')
  )

AfricaPlot
```

# Observed Elephant Populations in 28 African Countries 

```{r, echo=FALSE, eval=TRUE}
library(ggplot2)
library(readr)
library(dplyr)

d3 = african_elephants

d3 <- d3 %>%
  filter(
    Year == '2015',
    `African Elephant population (AfESG, 2019)` != '415428', 
    `African Elephant population (AfESG, 2019)` != '24119', 
    `African Elephant population (AfESG, 2019)` != '86373', 
    `African Elephant population (AfESG, 2019)` != '293447', 
    `African Elephant population (AfESG, 2019)` != '11489'
    )
  
PleaseLord = d3

Africa2015Plot = 
  PleaseLord |>
  ggplot(aes(x = `African Elephant population (AfESG, 2019)`, y = Entity, fill = `African Elephant population (AfESG, 2019)`)) + 
  geom_bar(stat = 'identity', width = 0.8) + 
  ggtitle('African Elephant Populations in 28 Countries (2015)') + 
  labs(
    x = 'Elephant Population',
    y = "Country",
    fill = 'Approximate Population'
  ) + 
  scale_x_continuous(
    breaks = c(5000,25000,50000,75000,100000,125000),
    labels = c('5K', '25K', '50K', '75K', '100K', '125K'),
    expand = expansion(mult = c(0, 0.05))
  ) +
  theme_bw() + 
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = 'bold')
  )

Africa2015Plot
```

# Africa's Three Highest Elephant Populations 
```{r, echo=FALSE, eval=TRUE}
library(ggplot2)
AfrMostElephants = 
  african_elephants %>%
  filter(Entity %in% c('Botswana', 'Tanzania', 'Zimbabwe'))

AfrMostElephants <- AfrMostElephants %>%
  filter(Year != '2007')

BTZPlot = 
  AfrMostElephants |>
  ggplot(aes(x = `African Elephant population (AfESG, 2019)`, y = Entity, fill = `African Elephant population (AfESG, 2019)`)) + 
  geom_bar(stat = 'identity') + 
  ggtitle('Highest African Elephant Populations (2015)') + 
  labs(
    x = 'Elephant Population',
    y = "Country",
    fill = 'Approximate Population'
  ) + 
  scale_x_continuous(
    breaks = c(5000,25000,50000,75000,100000,125000),
    labels = c('5K', '25K', '50K', '75K', '100K', '125K')
  ) +
  theme_bw() + 
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = 'bold')
  )
  
BTZPlot
```

# A Comparison of African Elephant Populations 
```{r, echo=FALSE, eval=TRUE, fig.width=10, fig.height=6, warning=FALSE}
library(patchwork)
library(ggplot2)

d3 = african_elephants

d3 <- d3 %>%
  filter(
    Year == '2015',
    `African Elephant population (AfESG, 2019)` != '415428', 
    `African Elephant population (AfESG, 2019)` != '24119', 
    `African Elephant population (AfESG, 2019)` != '86373', 
    `African Elephant population (AfESG, 2019)` != '293447', 
    `African Elephant population (AfESG, 2019)` != '11489'
    )
  
PleaseLord = d3

Africa2015Plot = 
  PleaseLord |>
  ggplot(aes(x = `African Elephant population (AfESG, 2019)`, y = Entity)) + 
  geom_bar(stat = 'identity', width = 0.8) + 
  labs(
    x = 'Elephant Population',
    y = "Country",
    fill = 'Approximate Population'
  ) + 
  scale_x_continuous(
    breaks = c(5000,25000,50000,75000,100000,125000),
    labels = c('5K', '25K', '50K', '75K', '100K', '125K'),
    expand = expansion(mult = c(0, 0.05))
  ) +
  theme_bw() 

BTZPlot = 
  AfrMostElephants |>
  ggplot(aes(x = `African Elephant population (AfESG, 2019)`, y = Entity, fill = `African Elephant population (AfESG, 2019)`)) + 
  geom_bar(stat = 'identity') + 
  labs(
    x = 'Elephant Population',
    y = "Country",
    fill = 'Approximate Population'
  ) + 
  scale_x_continuous(
    breaks = c(5000,25000,50000,75000,100000,125000),
    labels = c('5K', '25K', '50K', '75K', '100K', '125K')
  ) +
  theme_bw() 

Africa2015Plot + BTZPlot + 
  plot_annotation(
    title = 'African Elephant Populations in 28 Countries. \n Top 3 Highest Elephant Populations in Africa.',
    theme = theme(plot.title = element_text(size = 12, face = 'bold')))

```



# Extra Credit

Special Easter Egg from PS5 key.Rmd: "I'm a very special easter egg in that I only appear in the markdown file. Did you find me? Email Hart for your 3/27 extra credit!"
